### 2021-10-09 paper/output/document/paper41.tex
fixed contradiction

If the circumstances do not hold and the goal is achieved, then the maxim is vacuously effective, since 
it is vacuously willed (as described above). While this scenario is counterintuitive, it is not very 
interesting for my purposes because, when the circumstances do not hold, a maxim is not applicable. It 
doesn't really make sense to evaluate a maxim when it's not supposed to be applied. The maxim ``When on Jupiter,
read a book to one-up your nemesis" is vacuously effective because it can never be disproven.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{fun}\isamarkupfalse%
\ not{\isacharunderscore}universalizable\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}maxim{\isasymRightarrow}s{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}not{\isacharunderscore}universalizable\ M\ s\ {\isacharequal}\ {\isacharparenleft}{\isasymTurnstile}\ {\isacharparenleft}{\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}p{\isachardot}\ W\ M\ p\ w{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}\isactrlbold {\isasymnot}\ {\isacharparenleft}E\ M\ s{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{The maxim willed by subject $s$ is not universalizable if, for all people $p$, if $p$ wills M, then 
$M$ is no longer effective for $s$.%
}%
\begin{isamarkuptext}%
This formalizes Korgsaard's definition of the practical contradiction
interpretation:  a maxim is not universalizable 
if, in the world where the maxim becomes the standard practice (i.e. everyone acts on the maxim), the
agent's attempt to use the maxim's act to achieve the maxim's goal is frustrated. In other words, if 
the maxim is universally willed (captured by applying a universal qunatifier and the will function 
to the maxim on the LHS), then it is no longer effective for the subject $s$ (RHS above).

One worry is the order of the universal quantifier and the $\vDash$ or derivability symbol. In this 
representation, the universal quantifier is tightly bound to the left hand side of the equation, so it 
only quantifies into the statement ``All p will maxim M at any world." We can imagine an alternate
representation that quantifies over the sentence $\vDash W M p \longrightarrow \sim E M s$. This is 
a sentence of type $s \longrightarrow bool$, since it accepts a subject ($p$) as input and returns 
a boolean (the result of $p$ substituted into the sentence.) Thus, applying the universal quantifier 
$\forall p$ to this sentence results in a boolean that tracks the universalizability (or lack thereof) 
of the maxim. I chose my representation because it tracks the intuitive meaning of the universalizability 
test better: ``IF (the maxim is universalized) THEN (it is no longer effective)".

It's not clear to me if there's any intuitive difference between these two representations, but the lemmas 
below (using simplified versions of the representations) show that my representation is weaker than 
the representation where the quantifier has larger scope. This seems promising—the representation in 
lemma test\_2 is likely too strong because the quantifier quantifies into $M a$ as well. It really 
doesn't seem like this should make a difference, since $p$ never appears on the RHS expression. Let's
see if Prof. Amin has insight.%

@@

@@

@@

@@

@@

@@

%:%262=204%:%
%:%263=205%:%
%:%264=206%:%
%:%265=207%:%
%:%266=208%:%
%:%267=209%:%
%:%268=210%:%
%:%269=211%:%
%:%270=212%:%
%:%271=213%:%
%:%272=214%:%
%:%273=215%:%
%:%274=216%:%
%:%275=217%:%
%:%276=218%:%
%:%277=219%:%
%:%278=220%:%
%:%279=221%:%
%:%280=222%:%
%:%281=223%:%
%:%282=224%:%

>>>>

If the circumstances do not hold and the goal is achieved, then the maxim is vacuously effective, since 
it is vacuously willed (as described above). While this scenario is counterintuitive, it is not very 
interesting for my purposes because, when the circumstances do not hold, a maxim is not applicable. It 
doesn't really make sense to evaluate a maxim when it's not supposed to be applied. The maxim ``When on Jupiter,
read a book to one-up your nemesis" is vacuously effective because it can never be disproven.%

@@

\isacommand{abbreviation}\isamarkupfalse%
\ universalized{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}maxim{\isasymRightarrow}s{\isasymRightarrow}t{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}universalized\ {\isasymequiv}\ {\isasymlambda}M\ s{\isachardot}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isacharparenleft}{\isasymforall}p{\isachardot}\ W\ M\ p\ w{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ not{\isacharunderscore}universalizable\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}maxim{\isasymRightarrow}s{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}not{\isacharunderscore}universalizable\ {\isasymequiv}\ {\isasymlambda}M\ s{\isachardot}\ {\isacharparenleft}{\isasymTurnstile}\ {\isacharparenleft}universalized\ M\ s\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}\isactrlbold {\isasymnot}\ {\isacharparenleft}E\ M\ s{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline

@@

\isamarkupcmt{The maxim willed by subject $s$ is not universalizable if, for all people $p$, if $p$ wills M, then 
$M$ is no longer effective for $s$.%
}%
\begin{isamarkuptext}%
Below is my first attempt at formalizing Korgsaard's definition of the practical contradiction
interpretation:  a maxim is not universalizable 
if, in the world where the maxim becomes the standard practice (i.e. everyone acts on the maxim), the
agent's attempt to use the maxim's act to achieve the maxim's goal is frustrated. In other words, if 
the maxim is universally willed (captured by applying a universal qunatifier and the will function 
to the maxim on the LHS), then it is no longer effective for the subject $s$ (RHS above).%
\isacommand{abbreviation}\isamarkupfalse%
\ prohibited{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}maxim{\isasymRightarrow}s{\isasymRightarrow}t{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}prohibited\ {\isasymequiv}\ {\isasymlambda}{\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s{\isachardot}\ O{\isacharbraceleft}\isactrlbold {\isasymnot}\ {\isacharparenleft}will\ {\isacharparenleft}c{\isacharcomma}a{\isacharcomma}\ g{\isacharparenright}\ s{\isacharparenright}\ {\isacharbar}\ c{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ FUL{\isadigit{0}}{\isacharcolon}{\isacharcolon}bool\ \isakeyword{where}\ {\isachardoublequoteopen}FUL{\isadigit{0}}\ {\isasymequiv}\ {\isasymforall}\ c\ a\ g\ s{\isachardot}\ not{\isacharunderscore}universalizable\ {\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s\ {\isasymlongrightarrow}\ {\isasymTurnstile}{\isacharparenleft}{\isacharparenleft}prohibited\ {\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isamarkupcmt{This representation of the Formula of Universal Law reads, ``For all circumstances, goals, acts, 
and subjects, if the maxim of the subject performing the act for the goal in the circumstances is not 
universalizable (as defined above), then, at all worlds, in those circumstances, the subject 
is prohibited (obligated not to) from willing the maxim.%
}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}FUL{\isadigit{0}}\ {\isasymlongrightarrow}\ False{\isachardoublequoteclose}%

@@

\isacommand{using}\isamarkupfalse%
\ O{\isacharunderscore}diamond\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ prod{\isachardot}simps{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ split{\isacharunderscore}conv\ \isacommand{by}\isamarkupfalse%
\ fastforce%

@@

\begin{isamarkuptext}%
FUL0 is not consistent, and sledgehammer is indeed able to show that it implies a contradiction. 
The axioms that sledgehammer used are O\_diamond, which is \isa{{\isasymTurnstile}{\isasymlambda}w{\isachardot}\ ob\ {\isacharquery}B\ {\isacharquery}A\ {\isasymlongrightarrow}\ {\isasymnot}\ {\isasymTurnstile}\isactrlbold {\isasymnot}\ {\isacharquery}B\isactrlbold {\isasymand}{\isacharquery}A}. This axiom captures 
the idea that an obligation can't contradict its context. This is particularly problematic if the goal or 
action of a maxim are equivalent to its circumstances. In other words, if the maxim has already been 
acted on or the goal has already been achieved, then prohibiting it is impossible. 
In any model that has at least one term, it is possible to construct a maxim where the circumstances, goal, 
and act (once a subject acts on it) are all that same term, resulting in a contradiction.

To get around this, I will exclude what I call ``badly formed maxims," which are those maxims sucht that the goal has already been 
achieved or the act has already been acted on. Under my formalization, such maxims are 
not well-formed. To understand why, we can return to Korsgaard's and O'Neill's interpretations of a maxim as a practical
guide to action. A maxim is a practical principle that guides how we behave in everyday life. A 
principle of the form ``When you are eating breakfast, eat breakfast in order to eat breakfast," is not 
practically relevant. No agent would ever need to act on such a principle. It's not necessarily contradictory
or prohibited, but it is the wrong kind of question to be asking yourself. It is not a 
well-formed maxim, so the categorical imperative does not apply to it.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ well{\isacharunderscore}formed{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}maxim{\isasymRightarrow}s{\isasymRightarrow}i{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}well{\isacharunderscore}formed\ {\isasymequiv}\ {\isasymlambda}{\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s\ w{\isachardot}\ {\isacharparenleft}{\isasymnot}\ {\isacharparenleft}\ {\isacharparenleft}c\ \isactrlbold {\isasymrightarrow}\ g{\isacharparenright}\ w{\isacharparenright}{\isacharparenright}\ {\isasymand}\ {\isacharparenleft}{\isasymnot}\ {\isacharparenleft}\ {\isacharparenleft}c\ \isactrlbold {\isasymrightarrow}\ a\ s{\isacharparenright}\ w{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{This abbreviation formalizes the well-formedness of a maxim for a subject. The goal cannot be 
already achieved in the circumstances and the subject cannot have already performed the act.%
}\isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ FUL\ \isakeyword{where}\ {\isachardoublequoteopen}FUL\ {\isasymequiv}\ {\isasymforall}M{\isacharcolon}{\isacharcolon}maxim{\isachardot}\ {\isasymforall}s{\isacharcolon}{\isacharcolon}s{\isachardot}\ {\isacharparenleft}{\isasymforall}w{\isachardot}\ well{\isacharunderscore}formed\ M\ s\ w{\isacharparenright}\ {\isasymlongrightarrow}\ {\isacharparenleft}not{\isacharunderscore}universalizable\ M\ s\ {\isasymlongrightarrow}\ {\isasymTurnstile}\ prohibited\ M\ s\ {\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{Let's try the exact same formalization of the FUL as above, except that it only applies to 
maxims that are well-formed at every world.%
}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}FUL{\isachardoublequoteclose}\isanewline
\ \ \isacommand{nitpick}\isamarkupfalse%
{\isacharbrackleft}user{\isacharunderscore}axioms{\isacharcomma}\ falsify{\isacharequal}true{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\isanewline
%
\isamarkupcmt{The FUL does not hold in DDL, because nitpick is able to find a model for my system in which it is 
false. If the FUL were already a theorem of the system, adding it wouldn't make the system any more 
powerful, so this is the desired result.
$\color{blue}$ Nitpick found a counterexample for card s = 1 and card i = 1:

  Skolem constants:
    a = ($\lambda x. \_$)($s_1$ := ($\lambda x. \_$)($i_1$ := False))
    c = ($\lambda x. \_$)($i_1$ := True)
    g = ($\lambda x. \_$)($i_1$ := False)
    $\lambda w$. p = ($\lambda x. \_$)($i_1$ := $s_1$)
    s = $s_1$ $\color{black}$%
}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\ FUL{\isacharcolon}FUL\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{nitpick}\isamarkupfalse%
{\isacharbrackleft}user{\isacharunderscore}axioms{\isacharcomma}\ falsify{\isacharequal}false{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ simp\isanewline
%
\isamarkupcmt{The system is still consistent! Nitpick is able to find a model in which all axioms are satisfied, 
so this version of the FUL is consistent.
$\color{blue}$ Nitpick found a model for card i = 1 and card s = 1:

  Empty assignment $\color{black}$%
}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
During the process of making FUL0 consistent, I used Isabelle to gain philosophical insights, 
especially about vacuous maxims. I used Nitpick and Sledgehammer to quickly test if a small tweak 
to FUL0 fixed the problem or if I was still able to derive a contradiction.  I then realized that if 
I defined the circumstances, act, and goal as constant, then FUL0 was indeed consistent. After some 
experiementation, Prof. Amin correctly pointed out that, as constants, these three entities were 
distinct. When merely quantifying over (c, a, g), all members of a tuple could be equivalent. Within
a minute, I could formalize this notion, add it to FUL0, and test if it solved the problem. The fact 
that it did spurred my philosophical insight about vacuous maxims.

The logic confirmed that certain kinds
of circumstance, act, goal tuples are too badly formed for the categorical imperative to logically 
apply to them. The realization of this subtle problem would have been incredibly difficult without 
computational tools. The syntax and typing of Isabelle/HOL forced me to bind the free-variable $M$
in the FUL in different ways and allowed me to quickly test many bindings. The discovery of this 
logical inconsistency then enabled a philosophical insight about which kinds of maxims make sense as 
practical principles. This is one way to do computational ethics: model a system in a logic, use 
computational tools to refine and debug the logic, and then use insights about the logic to derive 
insights about the ethical phenonema it is modelling. This procedure parallels the use of proofs in 
theoretical math to understand the world.%

@@

\begin{isamarkuptext}%
One potential problem with my formalization is that it does not hugly use the modal nature of the system. 
All of the properties that the FUL investigates hold at all worlds, in effect removing the modal nature 
of the system. This approach simplifies logical and therefore computational complexity, improving 
performance. On the other hand, it doesn't use the full expressivity of DDL. If I run into problems 
later on, one option is to tweak the FUL to use this expressivity.%

@@

%:%262=203%:%
%:%263=204%:%
%:%265=205%:%
%:%266=206%:%
%:%269=208%:%
%:%270=209%:%
%:%271=210%:%
%:%272=211%:%
%:%273=212%:%
%:%274=213%:%
%:%276=216%:%
%:%277=216%:%
%:%278=217%:%
%:%279=218%:%
%:%280=219%:%
%:%281=219%:%
%:%283=220%:%
%:%284=221%:%
%:%285=222%:%
%:%286=223%:%
%:%287=223%:%
%:%288=224%:%
%:%289=225%:%
%:%290=225%:%
%:%292=225%:%
%:%296=225%:%
%:%297=225%:%
%:%298=226%:%
%:%299=226%:%
%:%300=226%:%
%:%309=228%:%
%:%310=229%:%
%:%311=230%:%
%:%312=231%:%
%:%313=232%:%
%:%314=233%:%
%:%315=234%:%
%:%316=235%:%
%:%317=236%:%
%:%318=237%:%
%:%319=238%:%
%:%320=239%:%
%:%321=240%:%
%:%322=241%:%
%:%323=242%:%
%:%324=243%:%
%:%326=245%:%
%:%327=245%:%
%:%328=246%:%
%:%330=247%:%
%:%331=248%:%
%:%332=248%:%
%:%333=249%:%
%:%334=250%:%
%:%335=250%:%
%:%337=251%:%
%:%338=252%:%
%:%339=252%:%
%:%340=253%:%
%:%341=254%:%
%:%342=254%:%
%:%343=255%:%
%:%344=255%:%
%:%346=255%:%
%:%350=255%:%
%:%351=255%:%
%:%353=256%:%
%:%354=257%:%
%:%355=258%:%
%:%356=259%:%
%:%357=260%:%
%:%358=261%:%
%:%359=262%:%
%:%360=263%:%
%:%361=264%:%
%:%362=265%:%
%:%363=266%:%
%:%371=266%:%
%:%372=267%:%
%:%373=268%:%
%:%374=268%:%
%:%375=269%:%
%:%376=270%:%
%:%377=270%:%
%:%378=271%:%
%:%379=271%:%
%:%381=271%:%
%:%385=271%:%
%:%386=271%:%
%:%388=272%:%
%:%389=273%:%
%:%390=274%:%
%:%391=275%:%
%:%392=276%:%
%:%402=278%:%
%:%403=279%:%
%:%404=280%:%
%:%405=281%:%
%:%406=282%:%
%:%407=283%:%
%:%408=284%:%
%:%409=285%:%
%:%410=286%:%
%:%411=287%:%
%:%412=288%:%
%:%413=289%:%
%:%414=290%:%
%:%415=291%:%
%:%416=292%:%
%:%417=293%:%
%:%418=294%:%
%:%419=295%:%
%:%420=296%:%
%:%424=298%:%
%:%425=299%:%
%:%426=300%:%
%:%427=301%:%
%:%428=302%:%
